### 2017-02-01 AnswersToComments-finalreview.tex
changes to letter. Still need to check against paper.

\documentclass[letterpaper,12pt]{letter}
\usepackage[utf8]{inputenc}
\usepackage[spanish, mexico]{babel}
\usepackage{calligra}

\usepackage[top=1in, left=1.2in,right=1.2in]{geometry}
%\longindentation=0pt

\renewcommand*{\opening}[2]{\ifx\@empty\fromaddress
 \thispagestyle{firstpage}%
 {\raggedleft#2\par}%
 \else % home address
 \thispagestyle{empty}%
 {\raggedleft\begin{tabular}{l@{}}\ignorespaces
 \fromaddress \\*[2\parskip]%
 #2 \end{tabular}\par}%
 \fi
 \vspace{10\parskip}%
 {\raggedright \toname \\ \toaddress \par}%
 \vspace{2\parskip}%
 #1\par\nobreak}

%% --- Font

\renewcommand*{\familydefault}{\sfdefault}
\renewcommand*{\sfdefault}{pag} % Avant Garde

\usepackage{amssymb}
\def\labelitemii{$\blacktriangle$} % Windows no encuentra el símbolo solicitado
\renewcommand{\theenumii}{\alph{enumii}}

%% -- Logo
\usepackage{graphicx}

\author{Verónica Esther Arriola Ríos}

\begin{document}
\begin{letter}{\textbf{}}
\includegraphics[height=4cm, keepaspectratio=true]{escudo_UNAM.jpg}\vspace*{-4cm}

\address{Universidad Nacional Autónoma de México\\ Facultad de Ciencias,\\ Departamento de Matemáticas,\\ México, D.F.}
\telephone{5622 5426}

\opening{Dear Editors,}{\today}
\vspace{1cm}

Please find attached the pdf of our revised submission to the special issue on “Sensorimotor Contingencies for Cognitive Robotics”.

We would like to thank you and all reviewers for timely and informative reviews, and for the minor changes requested. In this final revised version we address these final comments. We detail below how we have addressed each reviewer point one by one.

We hope that the paper is now in a satisfactory state for publication, and look forward to hearing from you.

\vspace{1cm}

\signature{Verónica E. Arriola-Rios\\
Jeremy L Wyatt}

\closing{Yours sincerely,}

%%enclosure listing
%%\encl{}

\end{letter}

\textbf{Response to reviewers’ comments}

We would like to thank all the reviewers most sincerely for their effort in reading and commenting on our paper.  We now address their final remarks.

\textbf{Reviewer: 1}

{\em The authors have successfully modified the work according to the reviewers comments. In my opinion the work is ready to be published under one condition: equation 11 should be replaced with equation 23 from the appendix.}

Thanks to this observation we noticed an inconsistency between the text and the equations. We see that this inconsistency made it seem to the reader that equation 23 should appear where we used equation 11.  However, rather than fixing the inconsistency by swapping the equations we have chosen to address it by changing the text in the paragraph above equation 11. We believe that only equations for the constraints should be shown in this section, as it is from the constraint equations that all elese arises. So we show a constraint equation for each of the length, area and angle. To this end we also change Equation 13 (from concerning energy to the constraint that underpins the energy). We show how both energy and all derived forces are obtained from these constraint equations in the appendix.  We hope this makes the presentation clearer in the manner intended by the reviewer. Thanks for spotting this.

\textbf{Reviewer: 2}

{\em In this revision, the authors solved the major problems in previous manuscript, but there are several small problems:}

{\em 1, Equation (5) shows $F=m\frac{\partial p}{\partial t}$, but it seems that $F=m\frac{\partial^2 p}{\partial t^2}$ is more reasonable, since force equals mass multiplying acceleration. This problem remains in Equation (6) and the paragraph between Equation (5) and Equation (6).}

Sorry you are right, we mixed notation which was not used in the article where the momentum of the particle is defined as $p=mv$.  All the affected equations were modified to avoid this and keep $p$ as the position of the particle.

{\em 2, Equation (13) and Equation (14) are confusing. Is it a symbol conflict? Why does the derivative of $\vec{p}$ equal $\vec{p}$?}

It is a mathematical trick to force over damped systems of springs, which do not oscillate.  The text in this paragraph was modified to make this more explicit.  Also, we added a quick explanation about the differential equation of the form $f(x) = f'(x)$.

\end{document}